Leadership across cultures and contexts: an exploratory analysis

### 2020-05-03 leadership_across_cultures_contexts.Rmd
Changes in intro.

<!--  CONCLUSION PARAGRAPH/STATEMENT. -->
Much of the theoretical leadership literature across disciplines suggests relative uniformity in many qualities of community leaders. Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be universal qualities of political and community leaders. In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses, and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions. We also envision a curvilinear relationship between leadership status and supernatural abilities and qualities. Community leaders among small-scale hunter-gatherers may be seen as experts in the use of ritual and supernatural "tools" though they typically do not have a monopoly in their use. In chiefdoms and and in early prehistoric state societies political leaders are often seen as divine and have special supernatural qualities commoners do not. In contemporary WEIRD societies there is a trend of secularization among political leaders, generally, although religious and spiritual leaders do have substantial influence over social and moral norms.

@@

<!-- BENEFITS: DECISIONS IN THEIR INTERST -->
The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019]. 
<!-- BENEFITS: HIGH SOCIAL STATUS -->
Leadership positions, including formalized institutionalized roles and informal ephemeral roles, are also often if not universally associated with increased social status (as previously discussed). Leadership roles, through their association with increased social status, are widely associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014]. Leaders and high-status individuals are also more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004]. Across many industries and organizations leadership positions are associated with increased financial compensation [@tsui_income_2017]. High social status, material wealth, intelligence, and leadership status are also viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].
<!-- NEED BETTER CITATIONS ABOVE -->

<!-- BENFITS: MATING -->
Across diverse populations and organizations, community leaders, political leaders, and organization leaders tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success. In a phylogenetic multilevel meta-analysis of 288 reported statistical associations among nonindustrial societies, @von_rueden_mens_2016 found that most measures of reproductive success were positively associated with male social status, independent of subsistence type or status measure.

@@

<!-- COSTS: OPPURTUNITY COSTS -->
Evolutionary models of leadership generally assume the benefits of leadership outweigh the costs, for both leaders and followers. However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership, across disciplines. Group responsibilities necessarily impose opportunity costs on individual priorities, given leadership responsibilities require investments of time, energy, and resources that could be invested elsewhere. 
<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - SSS -->
Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010]. Devising and proposing compromises and optimal outcomes can be cognitively challenging and the solutions leaders propose are unlikely to be equally satisfactory for everyone involved. Therefore, leaders stand to become the targets of animosity. In the absence of state-based systems and institutionalized inequalities, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including overt physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017]. Physically formidable individuals with strong social networks are likely to be able to bear these costs more so than their counterparts. Larger body size and high social capital are qualities often associated with leaders in small-scale societies [@glowacki_leadership_2015]. With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude. In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.

<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - LSS -->
Conceptualizing leadership as a possible solution to problems associated with collective goods [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation. In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs. @rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in urban neighborhoods and neighborhood organizations in WEIRD settings, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits. The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination). @rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization. Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position. @rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US. Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics. Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.

@@

Evolutionary theoretical models and empirical studies often seek to integrate findings within a broad framework of human cultural diversity. The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions leaders serve is difficult to capture. Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations [e.g., @garfield_review_2019; @von_rueden_leadership_2015]. Here, we seek to provide the first empirical and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample of nonindustrial, non-WEIRD populations. We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework. We also aim to explore how the qualities and functions of leaders vary within populations across social contexts and across populations with different subsistence strategies.

In sum, we sought to empirically and systematically accomplish the following four goals: 1) identify variation in the *qualities of leaders* and the *functions they serve*. 2) Identify *features* of the qualities and functions of leaders (e.g., do particular qualities or functions tend to covary or cluster together?) in reference to theoretical models or novel interpretations. Such features might lend support to existing theories; for instance they might correspond to qualities associated with "prestige" or "dominance" [@henrich_evolution_2001] or functions associated with conflict or cooperation [@gavrilets_collective_2015; @glowacki_evolutionary_2017; @hooper_theory_2010]. 3) Identify variation in the *costs and benefits* of leadership for both leaders and followers. And 4) determine the extent to which the qualities and functions of leaders, and their costs and benefits, varied across social contexts within populations, varied across populations with different subsistence strategies, and varied by continental region.

@@

We estimated the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 tested three theoretical models against the ethnographic record. The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme. The eHRAF is a subset of the larger HRAF and is continually growing in scope.

Whereas @garfield_evolutionary_2019 documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven. We draw on the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967]. The PSF aims to mitigate Galton's problem by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions. It comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text.

The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership seven key words describing leaders or followers. For the current study, two authors coded each text record for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive. See @garfield_evolutionary_2019 for additional details on the ethnographic sample and methodology.

@@

ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders. ZG and KS then coded all text records for group structure.

@@

Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types.

To supplement results from the aforementioned methods, we also performed textual analyses and explored variation measures of biases stemming from the ethnographic record that may have impacted results.

@@

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Supplementary Figure \@ref(fig:subsisttree)). Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures. The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`. The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one variable in the study.

@@

## Variation in study variables by subsistence strategy, continental region, leader sex, and group structure

To investigate variation in functions, qualities, costs, and benefits, we fit a multiple logistic mixed effects regression model of each variable as a function of subsistence strategy, continental region, leader sex, and group structure, with random intercepts for author nested within culture. We then compared the AIC value of each model to a model with only the fixed and random intercepts, retaining `r nrow(multi_aic)` models that showed at least a two-point reduction in AIC value. We then created heatmaps that displayed the relative estimated marginal means for each level of these categorical explanatory variables for each outcome variable. See Figures S\@ref(fig:heatmapsubsis), S\@ref(fig:heatmapregion), S\@ref(fig:heatmapsex), and S\@ref(fig:heatmapgroups), and Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars).

@@

@@

Relationships between leadership and high status, including causality and dependence, are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019]. Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in small-scale warfare or social-political movements). @cheng_dominance_2019 and @cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways. Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans high social status may be a necessary but insufficient condition for achieving and maintaining positions of leadership. In our view, social context is critical to these discussions, as is often mentioned but also overlooked. Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses. The "low status" male who initiates an inter-group raid on a neighboring group may not be widely respected within his community nor be a candidate for community leadership, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p. 71] called the "intragenerational social hierarchy."

>>>>

<!--  CONCLUSION PARAGRAPH/STATEMENT. -->
Much of the theoretical leadership literature across disciplines suggests relative uniformity in many qualities of community leaders. Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be universal qualities of political and community leaders. In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions. We also envision a curvilinear relationship between leadership status and supernatural abilities and qualities. Community leaders among small-scale hunter-gatherers may be seen as experts in the use of ritual and supernatural "tools" though they typically do not have a monopoly in their use. In chiefdoms and and in early prehistoric state societies political leaders are often seen as divine and have special supernatural qualities commoners do not. In contemporary WEIRD societies there is a trend of secularization among political leaders, generally, although religious and spiritual leaders do have substantial influence over social and moral norms.

@@

<!-- BENEFITS: DECISIONS IN THEIR INTERST -->
The ability to influence group processes can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019]. 
<!-- BENEFITS: HIGH SOCIAL STATUS -->
Leadership positions, including formal institutionalized roles and informal ephemeral roles, are also often if not universally associated with increased social status (as previously discussed). Leadership roles, through their association with increased social status, are widely associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014]. Leaders and high-status individuals are also more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004]. Across many WEIRD industries and organizations leadership positions are associated with increased financial compensation [@tsui_income_2017]. High social status, material wealth, intelligence, and leadership status are also generally viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].
<!-- NEED BETTER CITATIONS ABOVE -->

<!-- BENFITS: MATING -->
Across diverse populations and organizations, community, organization, and political leaders, tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success. In a phylogenetic multilevel meta-analysis of 288 reported statistical associations among nonindustrial societies, @von_rueden_mens_2016 found that most measures of reproductive success were positively associated with male social status, independent of subsistence type or status measure.

@@

<!-- COSTS: OPPURTUNITY COSTS -->
Evolutionary models of leadership generally assume the benefits of leadership outweigh the costs, for both leaders and followers, on average. However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership, across disciplines. Group responsibilities necessarily impose opportunity costs on individual priorities, given leadership responsibilities require investments of time, energy, and resources. 
<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - SSS -->
Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010]. Devising and proposing compromises and optimal outcomes can be cognitively challenging and the solutions leaders propose are unlikely to be equally satisfactory for everyone involved. Leaders therefore are at risk of becoming the targets of animosity. In the absence of state-based systems and institutionalized inequalities, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including overt physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017]. Physically formidable individuals with strong social networks are likely to be able to bear these costs more so than their counterparts. Larger body size and high social capital are qualities often associated with leaders in small-scale societies [@glowacki_leadership_2015]. With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude. In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.

<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - LSS -->
Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation. In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs. @rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in WEIRD urban neighborhood organizations, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits. The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination). @rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization. Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position. @rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US. Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics. Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.

@@

Evolutionary theoretical models and studies often seek to integrate findings within a broad framework of human cultural diversity. The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions leaders serve is difficult to capture. Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations [e.g., @garfield_review_2019; @von_rueden_leadership_2015]. Here, we seek to provide the systematic and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample of nonindustrial, non-WEIRD populations. We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework. We also aim to explore how these dimensions of leadership vary within populations across social contexts and across populations with different subsistence strategies, as well as by leader sex and contintental region.

In sum, we sought to empirically and systematically accomplish the following four goals: 1) identify variation in the *qualities of leaders* and the *functions they serve*. 2) Identify *features* of the qualities and functions of leaders (e.g., do particular qualities or functions tend to covary or cluster together?) in reference to theoretical models or novel interpretations. Such features might lend support to existing theories; for instance they might correspond to qualities associated with "prestige" or "dominance" [@henrich_evolution_2001] or functions associated with conflict or cooperation [@gavrilets_collective_2015; @glowacki_evolutionary_2017; @hooper_theory_2010]. 3) Identify variation in the *costs and benefits* of leadership for both leaders and followers. And 4) determine the extent to which the qualities and functions of leaders, and their costs and benefits, varied across social contexts within populations, varied across populations with different subsistence strategies, and varied by leader sex or continental region.

@@

We estimated the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 tested three theoretical models against the ethnographic record. The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme. The eHRAF is a continually growing subset of the larger HRAF.

Whereas @garfield_evolutionary_2019 documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven. We draw on the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967]. The PSF aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions. It comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text.

The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers. For the current study, two authors coded each text record for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive. See @garfield_evolutionary_2019 for additional details on the ethnographic sample and methodology.

@@

ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders. ZG and KS coded all text records for group structure.

@@

Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types and other population-level measures.

To supplement results from the aforementioned methods, we also performed textual analyses and explored variation in measures of biases stemming from the ethnographic record that may have impacted results.

@@

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Figure S\@ref(fig:subsisttree)). Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures. The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`. The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one variable in the study.

@@

## Variation in study variables by group structure, subsistence strategy, continental region, and leader sex

To investigate variation in functions, qualities, costs, and benefits, we fit a multiple logistic mixed effects regression model of each variable as a function of group structure, subsistence strategy, continental region, and leader sex, with random intercepts for author nested within culture. We then compared the AIC value of each model to a model with only the fixed and random intercepts, retaining `r nrow(multi_aic)` models that showed at least a two-point reduction in AIC value. We then created heatmaps that displayed the relative estimated marginal means for each level of these categorical explanatory variables for each outcome variable. See Figures S\@ref(fig:heatmapsubsis), S\@ref(fig:heatmapregion), S\@ref(fig:heatmapsex), and S\@ref(fig:heatmapgroups), and Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars).

@@

WARFARE H&G Figure S\@ref(fig:heatmapsubsis).

@@

Relationships between leadership and high status, including causality and dependence, are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019]. Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in small-scale warfare or social-political movements). @cheng_dominance_2019 and @cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways. Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans high social status may be a necessary but insufficient condition for achieving and maintaining positions of leadership. In our view, social context is critical to these discussions [@wiessner_power_2010], which is an idea that is often mentioned but under-developed and under-investigated empirically. Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses. The "low status" male who initiates an inter-group raid on a neighboring group may not be widely respected within his community nor be a candidate for community leadership, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p. 71] called the "intragenerational social hierarchy."
